---
title: "Assignment 2"
author: "Aude Marêché"
date: "02/11/2018"
output: html_document
---

```{r}
`%>%` <- magrittr::`%>%`
```

```{r}
source("functions.R")
```

## Exercise 1

```{r, cache = T}
iris_subset_1 <- iris[c(89:94, 108:112),]
iris_subset_2 <- iris[88:114,]

permut_1 <- permutation_twogroups(iris_subset_1, "Sepal.Width", "Species", "versicolor",
"virginica", difference_in_means)
ggplot2::ggplot(tibble::as_tibble(permut_1["permuted"]), ggplot2::aes(permuted)) + ggplot2::geom_vline(ggplot2::aes(xintercept = permut_1[["observed"]])) + ggplot2::geom_histogram() + ggplot2::xlim(-0.7, 0.7)
ts_p_value_1 <- two_sided_p_value(permut_1[[2]], permut_1[["observed"]])

permut_2 <- permutation_twogroups(iris_subset_2, "Sepal.Width", "Species", "versicolor",
"virginica", difference_in_means)
ggplot2::ggplot(tibble::as_tibble(permut_2["permuted"]), ggplot2::aes(permuted)) + ggplot2::geom_vline(ggplot2::aes(xintercept = permut_2[["observed"]])) + ggplot2::geom_histogram() + ggplot2::xlim(-0.7, 0.7)
ts_p_value_2 <- two_sided_p_value(permut_2[[2]], permut_2[["observed"]])

print(ts_p_value_1)
print(ts_p_value_2)
```

The differences between the histograms and the positions observed test statistics are due to differences in the size of the subsets of the data and to differences in the values of the data themselves. The bigger subset gives us a histogram that is much closer to a smooth Gaussian than than the other one, which is flatter with more jagged edges.

The two p-values are above 0.05 and thus are considered to be non-conclusive according to standard procedure, even if the p-value of the second subset is close to being significant. Here we can once again observe that a bigger dataset is more likely to yield significant results.

We can see that with such small datasets, adding even a little bit of data can produce big changes in the shape and caracteristics of a distribution.

## Exercise 2

### Task A

```{r}
stress_shift_3dict <- dplyr::filter(stressshift::stress_shift_unamb, Dict %in% c("W1802", "J1917", "C1687"))
length(stress_shift_3dict[[1]])
```

### Task B

```{r}
stress_shift_3dict_using_pipe <- stressshift::stress_shift_unamb %>% dplyr::filter(Dict %in% c("W1802", "J1917", "C1687"))
identical(stress_shift_3dict, stress_shift_3dict_using_pipe)
```

### Task C

```{r}
stress_shift_3dict_nouns <- dplyr::filter(stress_shift_3dict, Category == "Noun")
stress_shift_3dict_verbs <- dplyr::filter(stress_shift_3dict, Category == "Verb")
stress_shift_3dict_using_bind <- dplyr::bind_rows(stress_shift_3dict_nouns, stress_shift_3dict_verbs)
stress_shift_3dict_using_bind_reversed <- dplyr::bind_rows(stress_shift_3dict_verbs, stress_shift_3dict_nouns)
identical(stress_shift_3dict, stress_shift_3dict_using_bind)
identical(stress_shift_3dict, stress_shift_3dict_using_bind_reversed)
```

Only one of the tables created is identical to the original one because, while the values are the same in all tables, in a variable of type "table" (here actually a type "list") the values are considered to be ordered, so two tables who have the same values in different orders are not considered to be identical even if the values the tables contain are the same. This does not matter in a statistical analysis since we are only interested in the actual values and not in the order in which they are presented.

### Task D

```{r}
stress_shift_nouns_renamed <- dplyr::rename(dplyr::select(dplyr::filter(stressshift::stress_shift_unamb, Category == "Noun"), Word, Dict, Syllable), Syllable_Noun = Syllable)
stress_shift_verbs_renamed <- dplyr::rename(dplyr::select(dplyr::filter(stressshift::stress_shift_unamb, Category == "Verb"), Word, Dict, Syllable), Syllable_Verb = Syllable)
stress_shift_wide <- dplyr::inner_join(stress_shift_nouns_renamed, stress_shift_verbs_renamed)
```

The table `stress_shift_wide` now contains, for every word in every dictionary, if the stress shift occurs on the first syllable or the second syllable depending on if the word is a noun or a verb. This table has less rows than the sum of the two original tables because the values from these tables have been combined ("inner_joined") by word (and dictionary), which took out the need for separate rows to differentiate between noun and verb. This also left out words that were only nouns or only verbs, for which part of the information was missing.

### Task E

```{r}
ggplot2::ggplot(stressshift::stress_shift_unamb, ggplot2::aes(Category, fill = Syllable, stat = "identity")) + ggplot2::geom_bar(position = "dodge")
```

### Task F

```{r}
stress_shift_byword <- dplyr::summarize(dplyr::group_by(stress_shift_wide, Word),
                                        Noun_Percent_Syll_1 = (sum(Syllable_Noun == "Syllable 1") / n()) * 100,
                                        Verb_Percent_Syll_1 = (sum(Syllable_Verb == "Syllable 1") / n()) * 100)
length(stress_shift_byword[[1]])
```

### Task G

```{r}
ggplot2::ggplot(stress_shift_byword, ggplot2::aes(Noun_Percent_Syll_1, Verb_Percent_Syll_1)) + ggplot2::geom_point()
```

### Task H

```{r}
stress_shift_wide_all <- dplyr::full_join(stress_shift_nouns_renamed, stress_shift_verbs_renamed)
stress_shift_byword_all <- dplyr::summarize(dplyr::group_by(stress_shift_wide_all, Word),
                                        Noun_Percent_Syll_1 = (sum(Syllable_Noun == "Syllable 1", na.rm = TRUE) / sum(is.na(Syllable_Noun) == FALSE, na.rm = TRUE)) * 100,
                                        Verb_Percent_Syll_1 = (sum(Syllable_Verb == "Syllable 1", na.rm = TRUE) / sum(is.na(Syllable_Verb) == FALSE, na.rm = TRUE)) * 100)
```